---
title: "Problem Set 2"
output: html_notebook
---

This problem set covers concepts from chapters 2.2-2.5 in *Humanities Data in R*, including:

* Character Vectors
* Data Frames
* Matrices
* I/O
* Advanced Subsetting

As well as these concepts not covered in the book:

* For loops
* If statements

# Writing Code
For each of the following questions, write the code you need to solve the problem in the provided code block. Running each code block should result in a returned value-- either the result of a calcuation, or the printing of a variable. 

For example, rather than writing:
```{r}
x <- 1
```
you should write:
```{r}
x <- 1
x
```
Notice the difference between running these two code blocks.

Before submitting, be sure to run your code to make sure it is correct!

### Problems
1. Create a character vector called `rainbow` containing the names of six colors.
```{r}

```

2. Return the class of the vector `rainbow`.
```{r}

```

3. Return each color followed by the word "socks," separated by a space (i.e. "red socks").
```{r}

```

4. Return the first letter of each color.
```{r}

```

5. Create a vector with the full name of each of the days of the week, then use it to return the three-letter abbreviation of each.
```{r}

```

6. Use `grep` to find the index of each day that contains a capital "T". If you want to try something harder, use a regular expression (`?regexp`) to find the index of every day containing either a capital or lowercase "T".
```{r}

```

7. Make the code below function by casting the character vector as a numeric one.
```{r}
numbers <- c("1", "2", "3")
sum(numbers)
```


#### NYT Best Sellers List
The data in this section is taken from the New York Times Best Sellers combined print & e-book fiction list from the week of January 18, 2020. You can see where this data was drawn from online [here](https://www.nytimes.com/books/best-sellers/2020/01/19/combined-print-and-e-book-fiction/). There are fifteen books on the list. The csv has the following for each book: its number on the list; the title; the author; the number of weeks it has been on the list; and whether it has moved up, down, or not at all since last week.

8. Check your working directory, and set it to the assignment folder (if it is not already!). Then return the contents of the folder using `dir`.
```{r}

```

9. Load the file `nyt-best-sellers-jan-18.csv` into a data frame called `books`.
```{r}

```

10. Return the column names of `books`
```{r}

```

11. Find the average number of weeks on the list. Use data frame subsetting to select the appropriate column, and use `length()` to get the number of entries (don't hard-code in 15).
```{r}

```

12. Return the class of the `Movement` column.
```{r}

```


13. Notice in the Movement column that "Up" varies in its capitalization. Replace all instances of "up" with "Up" This might be tricky-- think "find and replace"! How do you find every instance of something? Then, how do you replace things?
```{r}


```


14. Now that we've normalized our values, we're going to modify this dataframe to make the Movement data more operation-friendly. Instead of one column with the words "up", "down", and "none", we want two columns with logical values: one for whether a book moved up, and one for whether a book moved down. 

a. Create a new column in `books` called "MovedUp" that contains logical values (i.e. TRUE and FALSE) stating whether the book moved up that week.
```{r}

```

b. Do the same for "MovedDown".
```{r}

```

c. Remove the Movement column
```{r}

```

15. For whichever book you please, use the dataframe to print a sentence like "Book Title by Author Name has been on the best sellers list for Number weeks. It is currently number X."
```{r}

```

16. Return the subset of `books` that have moved up this week. 
```{r}

```

17. Complete the following code to print the title and author of each book that has moved up this week. For each book, print something like "Title by Author"
```{r}
print("Books moving up:")
for( ){
  if( ){
    print(paste( ))
  }
}

```

18. Print a string like "Title by Author stayed at number X" for every book that *was* on the list last week but did *not* move. As a sanity check: you should get 5 results.
```{r}


```

19. Save the changes you've made to the data. Write `books` to a file called "nyt-list-modified.csv". Since we already have a Number column, add `row.names=FALSE` as an additional parameter (after the filename in your code) to stop R from adding another column of row numbers. Submit this file along with your R notebook.
```{r}

```


# Explaining Code
For each of the following questions, explain what's happening in the code provided *without* adding anything to the code. Answer all parts of the question.

20. Why does this return an error?
```
"1" + "2"
```

21. Write the following code in English. That is, how would you read this code out loud if you were explaining what it does?
```
names <-  c("Asa", "Lena", "Mia", "Daniel", "Zahara")
names <-  sort(names)
for(i in 1:length(names)){
  print(substr(names[i], start=1, stop = 1))
}
```
